---
title: "CPC Merged Analysis of Precipitation (CMAP)"
output: html_document
date: '2022-05-30'
---

```{r}
library(ncdf4)

# set path and filename
ncpath <- "/Users/hwingren/Downloads/Blackwell/Blackwell_Scholars_2022/"
ncname <- "precip.mon.mean.nc"  
ncfname <- paste(ncpath, ncname, sep="")
dname <- "precip"
```


# open a netCDF file
```{r}
ncin <- nc_open(ncfname)
print(ncin)
```

# get longitude and latitude
```{r}
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"lat")
nlat <- dim(lat)
head(lat)

print(c(nlon,nlat))
```

# get time
```{r}
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt

tunits
```

# get precipitation
```{r}
precip_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(precip_array)
```

# get global attributes
```{r}
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

nc_close(ncin)
```

# load some packages
```{r}
library(chron)
library(lattice)
library(RColorBrewer)
```

# convert time -- split the time units string into fields
```{r}
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])
chron(time,origin=c(tmonth, tday, tyear))
```

# replace netCDF fill values with NA's
```{r}
precip_array[precip_array==fillvalue$value] <- NA

length(na.omit(as.vector(precip_array[,,1])))
```

# get 12 months slice of layer (Jan-Dec 1982)
```{r}
#m <- 1
precip_slice <- precip_array[,,c(37:48, 241:252)]
dim(precip_slice)
```

# quick map
```{r}
for(m in 1:24){
image(lon,rev(lat),precip_slice[, ncol(precip_slice):1,m], col=rev(heat.colors(16)), breaks=0:16)}
```

# levelplot of the slice
```{r}
grid <- expand.grid(lon=lon, lat=lat)
cutpts <- seq(0, 40, length.out=11)
m<-1
levelplot(precip_slice[,,m] ~ lon * lat, data=grid, at=cutpts, cuts=11, pretty=T, 
  col.regions=(rev(heat.colors(16))))
```
            
# create new lon and lat vectors
```{r}
lon_new <- c(lon[73:144]-360, lon[1:72])
lat_new <- rev(lat)
```

# create new precipitation matrix
```{r}
precip_slice_new <- precip_slice[c(73:144, 1:72), ncol(precip_slice):1,]
```

# plot whole globe
```{r}
m<-1 #January
image(lon_new, lat_new, precip_slice_new[,,m], col=rev(heat.colors(16)), breaks=0:16)
#Compare with previous plot
image(lon,rev(lat),precip_slice[, ncol(precip_slice):1,m], col=rev(heat.colors(16)), breaks=0:16)
```

# create boolean masks corresponding to region of interest
```{r}
lon_mask <- -75<=lon_new & lon_new <=(-60) # need parentheses here because <- is the assignment operator
lat_mask <- 0<=lat_new & lat_new<=12
```

# plot region of interest
```{r}
m<-1
for(m in 1:24){
image(lon_new[lon_mask], lat_new[lat_mask], precip_slice_new[lon_mask, lat_mask,m], col=rev(heat.colors(16)), breaks=0:16)} # I took off the rev function here
#
m<-24
grid <- expand.grid(lon=lon_new[lon_mask], lat=lat_new[lat_mask])
grid$z<-as.vector(precip_slice_new[lon_mask, lat_mask,m])
cutpts <- seq(0, 16, length.out=11) #watch out for the limits here
levelplot(z ~ lon * lat, data=grid, at=cutpts, cuts=11, pretty=T, 
  col.regions=(rev(heat.colors(16))))
```

# create dataframe -- reshape data
```{r}
# matrix (nlon*nlat rows by 2 cols) of lons and lats
lonlat <- as.matrix(expand.grid(lon_new[lon_mask],lat_new[lat_mask]))
dim(lonlat)
```

# vector of `precip` values
```{r}
prec_mat<-lonlat
for (m in 1:24) {
precip_vec <- as.vector(as.vector(precip_slice_new[lon_mask, lat_mask,m]))
prec_mat<-cbind(prec_mat,precip_vec)
}
colnames(prec_mat) <- c("lon","lat","Jan_1982","Feb_1982","Mar_1982","Apr_1982","May_1982","Jun_1982",
  "Jul_1982","Aug_1982","Sep_1982","Oct_1982","Nov_1982","Dec_1982","Jan_1999","Feb_1999","Mar_1999","Apr_1999","May_1999","Jun_1999",
  "Jul_1999","Aug_1999","Sep_1999","Oct_1999","Nov_1999","Dec_1999")
```

```{r}
mat_stat<-apply(prec_mat[,3:26],2,summary)
mat_stat
```

# calculating statistics for full year
```{r}
summary(as.vector(prec_mat[, 3:14]))
summary(as.vector(prec_mat[, 15:26]))
sd(as.vector(prec_mat[, 3:14]))
sd(as.vector(prec_mat[, 15:26]))
```