---
title: "CPC Merged Analysis of Precipitation (CMAP)"
output:
  pdf_document: default
  html_document: default
date: '2022-05-30'
---

```{r}
library(ncdf4)

# set path and filename
ncpath <- "/Users/hwingren/Downloads/Blackwell/Blackwell_Scholars_2022/"
ncname <- "precip.mon.mean.nc"  
ncfname <- paste(ncpath, ncname, sep="")
dname <- "precip"
```


# open a netCDF file
```{r}
ncin <- nc_open(ncfname)
print(ncin)
```

# get longitude and latitude
```{r}
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon)
head(lon)

lat <- ncvar_get(ncin,"lat")
nlat <- dim(lat)
head(lat)

print(c(nlon,nlat))
```

# get time
```{r}
time <- ncvar_get(ncin,"time")
time

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time)
nt

tunits
```

# get precipitation
```{r}
precip_array <- ncvar_get(ncin,dname)
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(precip_array)
```

# get global attributes
```{r}
title <- ncatt_get(ncin,0,"title")
institution <- ncatt_get(ncin,0,"institution")
datasource <- ncatt_get(ncin,0,"source")
references <- ncatt_get(ncin,0,"references")
history <- ncatt_get(ncin,0,"history")
Conventions <- ncatt_get(ncin,0,"Conventions")

nc_close(ncin)
```

# load some packages
```{r}
library(chron)
library(lattice)
library(RColorBrewer)
```

# convert time -- split the time units string into fields
```{r}
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])
chron(time,origin=c(tmonth, tday, tyear))
```

# replace netCDF fill values with NA's
```{r}
precip_array[precip_array==fillvalue$value] <- NA

length(na.omit(as.vector(precip_array[,,1])))
```

# get 12 months slice of layer (Jan-Dec 1982 and Jan-Dec 1999)
```{r}
#m <- 1
precip_slice <- precip_array[,,c(37:48, 241:252)]
dim(precip_slice)
```

# quick map
```{r}
for(m in 1:24){
image(lon,rev(lat),precip_slice[, ncol(precip_slice):1,m], col=rev(heat.colors(16)), breaks=0:16)}
```

# levelplot of the slice
```{r}
grid <- expand.grid(lon=lon, lat=lat)
cutpts <- seq(0, 40, length.out=11)
m<-1
levelplot(precip_slice[,,m] ~ lon * lat, data=grid, at=cutpts, cuts=11, pretty=T, 
  col.regions=(rev(heat.colors(16))))
```
            
# create new lon and lat vectors
```{r}
lon_new <- c(lon[73:144]-360, lon[1:72])
lat_new <- rev(lat)
```

# create new precipitation matrix
```{r}
precip_slice_new <- precip_slice[c(73:144, 1:72), ncol(precip_slice):1,]
```

# plot whole globe
```{r}
m<-1 #January
image(lon_new, lat_new, precip_slice_new[,,m], col=rev(heat.colors(16)), breaks=0:16)
#Compare with previous plot
image(lon,rev(lat),precip_slice[, ncol(precip_slice):1,m], col=rev(heat.colors(16)), breaks=0:16)
```

# create boolean masks corresponding to region of interest
```{r}
lon_mask <- -75<=lon_new & lon_new <=(-60) # need parentheses here because <- is the assignment operator
lat_mask <- 0<=lat_new & lat_new<=12
```

# plot region of interest
```{r}
m<-1
for(m in 1:24){
image(lon_new[lon_mask], lat_new[lat_mask], precip_slice_new[lon_mask, lat_mask,m], col=rev(heat.colors(16)), breaks=0:16)} # I took off the rev function here
#
m<-24
grid <- expand.grid(lon=lon_new[lon_mask], lat=lat_new[lat_mask])
grid$z<-as.vector(precip_slice_new[lon_mask, lat_mask,m])
cutpts <- seq(0, 16, length.out=11) #watch out for the limits here
levelplot(z ~ lon * lat, data=grid, at=cutpts, cuts=11, pretty=T, 
  col.regions=(rev(heat.colors(16))))
```

# create dataframe -- reshape data
```{r}
# matrix (nlon*nlat rows by 2 cols) of lons and lats
lonlat <- as.matrix(expand.grid(lon_new[lon_mask],lat_new[lat_mask]))
dim(lonlat)
```

# vector of `precip` values
```{r}
prec_mat<-lonlat
for (m in 1:24) {
precip_vec <- as.vector(as.vector(precip_slice_new[lon_mask, lat_mask,m]))
prec_mat<-cbind(prec_mat,precip_vec)
}
colnames(prec_mat) <- c("lon","lat","Jan_1982","Feb_1982","Mar_1982","Apr_1982","May_1982","Jun_1982",
  "Jul_1982","Aug_1982","Sep_1982","Oct_1982","Nov_1982","Dec_1982","Jan_1999","Feb_1999","Mar_1999","Apr_1999","May_1999","Jun_1999",
  "Jul_1999","Aug_1999","Sep_1999","Oct_1999","Nov_1999","Dec_1999")
```

```{r}
mat_stat<-apply(prec_mat[,3:26],2,summary)
mat_stat
x = prec_mat[,1]
y = prec_mat[,2]
trend_one = lm(prec_mat[,3]~x+y)
trend_two = lm(prec_mat[,3]~x+y+I(x^2)+I(y^2)+I(x*y))
summary(trend_one)
summary(trend_two)
```

# calculating statistics for full year
```{r}
summary(as.vector(prec_mat[, 3:14]))
summary(as.vector(prec_mat[, 15:26]))
sd(as.vector(prec_mat[, 3:14]))
sd(as.vector(prec_mat[, 15:26]))
```

# calculate an empirical semi-variogram
```{r}
library(geoR)
geo_precip_list = list()
empirical = list()
for(m in 3:26) {
    atrend="1st"
    geo_precip = as.geodata(prec_mat, coords.col = 1:2, data.col = m)
    geo_precip_list[[m-2]] = geo_precip
    if (m == 24) atrend="2nd"
    empirical[[m-2]] = variog(geo_precip, trend = atrend)
    plot(empirical[[m-2]])
}

```

# fitting a model
```{r}
for(m in 1:24) {
    if(m %in% c(3, 4, 6, 10, 14, 19, 20, 23, 24)) {
        fit = variofit(empirical[[m]], cov.model = "gaussian", fix.nugget = FALSE, fix.kappa = TRUE)
        my_title = paste0("Gaussian Variogram for Month ", m)
    } else if(m %in% c(1, 2, 5, 7, 8, 9, 11, 12, 13, 15, 16, 17)) {
        fit = variofit(empirical[[m]], cov.model = "spherical", fix.nugget = FALSE, fix.kappa = TRUE)
        my_title = paste0("Spherical Variogram for Month ", m)
    } else {
        fit = variofit(empirical[[m]], cov.model = "exponential", fix.nugget = FALSE, fix.kappa = TRUE)
        my_title = paste0("Exponential Variogram for Month ", m)
    }
    plot(empirical[[m]], main=my_title)
    lines(fit)
}
fit = variofit(empirical[[1]], cov.model = "spherical", fix.nugget = FALSE, fix.kappa = TRUE)

#Months 18, 21, all 4 equal
#Month 22, matern and exponential equal
#gaussian = 3, 4, 6, 10, 14, 19, 20, 23, 24
#spherical = 1, 2, 5, 7, 8, 9, 11, 12, 13, 15, 16, 17
```

#Load Libraries
```{r}
library(atakrig)
library(raster)
library(terra)
```

#Calculate Residuals
```{r}
resid_mat <- prec_mat[, 1:2]
for(m in 1:24) {
    mod <- lm(prec_mat[, 2+m] ~ prec_mat[, 1] + prec_mat[, 2])
    resid_mat <- cbind(resid_mat, resid(mod))
}
```

#Discretize raster to points
```{r}
deconv_precip<-list()
for(m in 1:24) {
  prec_ras_resid = raster(apply(matrix(resid_mat[,m], nrow=5, ncol=6, byrow=TRUE), 2, rev), xmn=-73.75, xmx=-61.25, ymn=1.25, ymx=11.25)
  obs.discrete = discretizeRaster(prec_ras_resid, cellsize=480, psf = "gau", sigma = 2)
    if(m %in% c(3, 4, 6, 10, 14, 19, 20, 23, 24)) {
      deconv_precip[[m]] = deconvPointVgm(obs.discrete, model = "Gau", maxIter = 100, fig=TRUE, ngroup = 24, longlat = FALSE)
    } else if(m %in% c(1, 2, 5, 7, 8, 9, 11, 12, 13, 15, 16, 17)) {
        deconv_precip[[m]] = deconvPointVgm(obs.discrete, model = "Sph", maxIter = 100, fig=TRUE, ngroup = 24, longlat = FALSE)
    } else {
        deconv_precip[[m]] = deconvPointVgm(obs.discrete, model = "Exp", maxIter = 100, fig=TRUE, ngroup = 24, longlat = FALSE)
    }
}
```

#Make Kriging predictions
```{r}
new_lonlat = expand.grid(x = seq(-73.75,-61.25,length.out = 24), y = seq(1.25,11.25,length.out = 20))

for(m in 3:26) {
  krige.control(type.krige = "ok", trend.d = "1st", trend.l = "1st",
            obj.model = NULL, cov.model = "spherical", cov.pars = deconv_precip[[m]]$pointVariogram[,2:3])
  krige.conv(geo_precip_list[[m-2]], coords=geo_precip_list[[m-2]]$coords, data=geo_precip_list[[m-2]]$data,
           locations = new_lonlat, krige = krige.control, output)
}
```